---
title: "Functions to help with multilevel modelling using DataSHIELD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-information}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dsHelper)
library(MolgenisArmadillo)
library(DSMolgenisArmadillo)
library(tidyverse)
library(dsHelper)
```
```{r, eval = FALSE, echo = FALSE}
setwd("C:/Users/timca/OneDrive - University of Bristol/repos")
rmarkdown::render("./ds-helper/vignettes/multi-level-modelling.rmd", output_file="multi-level-modelling.html")
```

## Overview
Multi-level/mixed effect modelling is now possible within DataSHIELD. There are 
number of step to go through when fitting multilevel models (e.g. QC, diagnostics, 
model selection) and this isn't quite as straightforward. This vignette shows how
to use a suite of functions I've created when fitting multilevel models.

[See also the tutorial I will write on multilevel modelling using datashield, 
also I've found out a WP7 deliverable]

## Accessing the data for this vignette
To access the example data used in this vignette [explanation by Sido + link to
login vignettes]
